A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-03-06 paper/paper1/ms.tex
experimenting with p-winds to explore mass loss rates

As mentioned in Section \ref{pwinds}, inferring the mass loss rate from observations is strongly model dependent, as seen in \censorbox{cite e.g. S.V. papers}.  In particular, the uncertain XUV-to-NUV flux ratio introduces a large degeneracy in

>>>>

As mentioned in Section \ref{pwinds}, inferring the mass loss rate from observations is strongly model dependent, as seen in \censorbox{cite e.g. S.V. papers}.  The XUV flux, exosphere temperature $T_0$, and $\dot{M}$ are all at least partially degenerate.  In particular, the uncertain XUV-to-NUV flux ratio confounds the interpretation of the observed Helium signal.  Even the ``high XUV'' scenario in Figure \ref{fig:XUV} does not have enough ground-state ionizing photons to penetrate to within 5 $R_p$ of the planet surface.  This means that at best only 1 in $10^{6}$ Helium atoms participate in the observability of the Helium metastable triplet lines in the baseline 1D Parker wind model.  Turning down the XUV flux makes the signal even less observable.  A larger mass loss rate would be needed to explain the same observed HPF Helium absorption, if say only 1 in $10^{7}$ helium atoms are populating the metastable state.

Finally, the entire premise of the Parker wind model may be unjustified due to the inherently 3D geometry of the leading tail.  In the next section we critically examine the assumptions in the 1D Parker wind model.

\subsection{Critical examination of the Parker Wind model}
An even more existential limit arises when we consider the assumptions of the isothermal Parker wind model, and the extreme geometry of \hatpb.  Figure \ref{fig:murrayClaySchematic} shows a schematic modeled after Figure 8 from \citet{2009ApJ...693...23M}.  The point at which the gas velocity exceeds the sound speed, the so-called sonic point, resides at 1.2 $R_p$, a mere $\sim9$ atmospheric scale heights from the planet surface.  The smallness of this radius can be understood from the significant effect of tidal gravity and the .  The pressure level here would be in the fractions of a millibar level, with an atmospheric temperature $\le T_\mathrm{eq}\sim10^3$~K and not yet the $T\sim10^4$~K exosphere range.  Depending on the mass loss rate, the transition from primarily neutral to primarily ionized plasma may occur outside or inside the sonic radius.

\begin{figure}
    \includegraphics[width=\linewidth]{figures/murray_clay_schematic_v0p1.png}
    \caption{Schematic of the relevant length scales in a stratified atmosphere, modeled after Figure 8 from \citet{2009ApJ...693...23M}.  The sonic point resides at a mere 2 atmospheric scale heights, and well below the photoionization level of Hydrogen, raising the possibility that the supersonic flow begins neutral and becomes ionized later on.}
    \label{fig:murrayClaySchematic}
\end{figure}

Eventually, for very large mass loss rates, the width of the line begins to grow.
